\chapter{Summary}

In this thesis I study the problem of detecting fake reviews written by the same users under different identities and aim to use semantic similarity metrics to compute the relatedness between user reviews. This represents a new approach to opinion spam detection, meant to capture more information than a simple text similarity measure, such as the cosine similarity and tap deeper into the textual particularities of a review. It is aimed at spammers which use multiple anonymous profiles to post reviews by reusing previous text they had also written, replacing the main aspect words with synonyms, while still keeping the overall sentiment the same. 

There is one obvious but important assumption - the imagination of a spammer, like any human is limited and thus will not be able to write completely different reviews about an imaginary experience every single time. Thus, it is very likely he will partially reuse some of the content and aspects between reviews.

This thesis proposes a complete solution to detect opinion spam of one-time reviewers using semantic similarity. It also proposes a method to detect opinion spam using recent research models aimed at extracting product aspects from short texts, based on topic modeling and in particular on Latent Dirichlet Allocation. Another goal was to test this hypothesis on real-life reviews and make a comparison with the existing cosine-based vectorial similarity models.

This research also proposes to use density-based clustering to group reviews using simple behavioral features, proven to be linked to suspicious user activity. Clustering divides the reviews into smaller chunks, thus the number of comparisons is reduced considerably. The novelty of the topic modeling method is to use the similarity of the underlying topic distributions of the reviews to classify them as truthful or deceptive. 

Experimental results showed that semantic similarity can outperform the vectorial model in detecting deceptive reviews, capturing even more subtle textual clues. The precision score of the review classifier showed high results, enough to make the method viable to be integrated into a production detection system. The results of the topic modeling method showed that combining opinion spam detection with topic modeling may offer, for some number of topics, results in the vicinity of the semantic similarity models.